---
title: "Aides Drug"
author: "Anthony Benitez"
date: "9/22/22"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->
I will be using my tricks helper file so i can change the character variable to factor variable and using Numerical, graphical, inferential methods so i can solve these problems and filter for only the drugusers.
```{r}
DrugUser<-filter(triUmp, DrugUser == "yes")
```

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results
```{r eval=FALSE}
triUmp <- triUmp %>%
mutate(across(where(is.character), as.factor))
```
### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->
```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")

```

There are 21743 people who took part in the test but there is only 10% of drug users in that population and from the looks of the bar chart there were 75% of people who took the placebo and of the people who got sick it was only 25% who took the drug .

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

Out of 100 percent of the people that got aids people who took the drugs only 24.29 percent got sick and the ones who took the placebo only 75.71 percent got sick. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The fishers test odds ratio shows us that if we took the drug we would have a 0.354128 chance of getting aids compared to someone taking a placebo.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
There are 21743 people who took part in the test but there is only 10% of drug users in that population and from the looks of the bar chart there were 75% of people who took the placebo and of the people who got sick it was only 25% who took the drug and out of 100 percent of the people that got aids people who took the drugs only 24.29 percent got sick and the ones who took the placebo only 75.71 percent got sick. The fishers test odds ratio shows us that if we took the drug we would have a 0.354128 chance of getting aids compared to someone taking a placebo. With all of our results that we got the drug does work and the odds of getting aids is 0.354. These results only applicable to drug users.
